Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
miscellaneous edits

\begin{abstract}

@@

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence.  Recent work in program induction show
that deep networks can be very successful at learning to write certain
kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
techniques (e.g., Metagol:~\cite{muggleton2015meta} and 
FlashFill:~\cite{gulwani2011automating}).
However, the success of these approaches -- both symbolic and neural --
relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
The DSL is an inventory of restricted programming primitives,
and imparts domain specific knowledge about the space of programs.
To what extent can we dispense with hand-engineered DSLs?

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to distill the commonalities across the programs that solve the tasks,
letting the agent solve other related program learning problems.

@@

  \end{tikzpicture}
  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). Observe tasks $x$ (program inputs/outputs). (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs
    ($\theta^{(x)} = q(x)$). Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of as a kind of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means  updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
An important difference from the Helmholtz machine
is that our setup has two distinct sleep phases: a phase where we update the generative model
$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
and a second phase where we update the recognition model $q$,
which we call \textbf{Sleep-R}.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternatingly maximizing w.r.t. the frontiers (``wake'') and the DSL (``sleep''):
%So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
Adding new programs to the frontiers means searching for new programs $p$ for task $x$
where $\probability[x,p|\mathcal{D},\theta]$ is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by learning a neural recognition model
during a second ``sleep'' phase:

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  With $q$ in hand we can find programs for
the frontier $\mathcal{F}_x$ by searching for programs that maximize
$q(p|x)$.
The network $q$ exploits the structure of the DSL $\mathcal{D}$:
rather than directly predicting a distribution over $p$ conditioned on $x$,
it predicts a weight vector, $\theta^{(x)}$, and we define $q(p|x)\triangleq \probability[p|\mathcal{D},\theta = q(x)]$.
This taps into the intuition that programming is primarily a top-down activity:
as human programmers, we first decide what kind of
programming constructs we might need to use,
and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Figure~\ref{graphicalModel}.

Our lower bound $\lowerBound$ is unconventional,
and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
For the interested reader, we explain in the supplement that maximizing an ELBO-style bound
leads to a pathological behavior that would cause the algorithm to easily become trapped in local optima.

@@

The recognition model $q$ is a neural network that predicts,
for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
Together with the DSL, this defines a distribution over programs,
$\probability[p|\mathcal{D},\theta = q(x)]$.
We abbreviate this distribution as $q(p|x)$.
The crucial aspect of this framing is that the neural network
leverages the structure of the learned DSL,
so it is \emph{not} responsible for
generating programs wholesale.

@@

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Appendix we derive a tractable MAP estimator for $\theta$.
\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) on the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a DSL has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
(2) Because the frontiers can grow very large,
we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(3) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
(4) For added robustness, at each iteration we enumerate programs from both the generative model and the recognition model.

@@

This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
%% For this domain our recognition model is a simple multilayer perceptron 
%% whose inputs are the input/output examples.
\system learns to build these
circuits by creating new subroutines that correspond to
logic gates and multiplexers (Figure~\ref{multiplexer}).
We will show data showing that the neural recognition model
assists in building these circuits (Section~\ref{quantitative}),
suggesting that an agent can glean something about the structure of the circuit just by looking at its truth table.

@@

We take approaches from two different initial DSLs, \emph{base} and
\emph{rich}, and assess their performance. The base DSL starts with only
low-level primitives such as \code{mapi}, \code{reducei}, and
\code{if}, whereas the rich DSL includes some common structure that can be
built from these such as \code{all}, \code{filter}, \code{slice}, and
\code{index}.
Both are theoretically capable of solving all tasks supplied by our dataset.
See Supplement for more detail on these DSLs.

@@

  \\\midrule\\\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  84\% &76\%& 63\%&62\%&62\%\\
  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s

  \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&    98\% &94\%&0.7\%&0\%&2\% \\
  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
  MDL (nats)&   9.3 &11.0&2.3&--&4.6

  \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&99\% &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
  \% solved&52\% &52\% &40\%&44\%&42\%\\
  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
  \% solved&86\% &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

@@

 \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
 focused on creating neural networks that regress from
 input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is analogous to the lesioned version of
 \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned (RF/DC baseline).
 These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
 In our work we consider a weakly-supervised setting where ground truth programs are not provided and
 the agent must learn from a few hundred examples.

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and go beyond it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.

@@

 \section{Conclusion}
 We contributed a model, \system, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the desired programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.

>>>>

\begin{abstract}
  Successful  approaches to program induction require a
  hand-engineered  domain-specific languaget (DSL),
  constraining the space of allowed programs and imparting
  prior knowledge of the domain.
  We contribute a program learning algorithm
  that learns the DSL called \system,
which infers
  a DSL 
  and trains a neural network to
  efficiently search for programs in the learned DSL.
  We use our model to build circuits,
  edit strings, do symbolic regression,
  and synthesize Lisp-style functions on lists,
  in each case showing that
  \system learns a domain-specific vocabulary for expressing programs.

@@

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence.  Recent work in program induction show
that deep networks can be very successful at learning to write certain
kinds of programs (e.g., RobustFill:~\cite{devlin2017robustfill} and
DeepCoder:~\cite{balog2016deepcoder}), as can symbolic search
techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}, and Sketch~\cite{solar2008program}).
However, the success of these approaches -- both symbolic and neural --
relies crucially upon a hand-engineered \emph{Domain Specific Language} (DSL).
The DSL is an inventory of restricted programming primitives,
and imparts domain specific knowledge about the space of programs.
To what extent can we dispense with hand-engineered DSLs?

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to expose the commonalities across the programs that solve tasks within a domain,
letting the agent solve other related program learning problems.

@@

  \end{tikzpicture}
  \caption{(a) DSL $\mathcal{D}$ generates programs $p$ by sampling DSL primitives with probabilities $\theta$ (Algorithm~\ref{programGenerativeModel}). Observe tasks $x$ (program inputs/outputs). (b)  Neural network $q(\cdot )$, the \emph{recognition model}, regresses from $x$ to the parameters of a distribution over programs,
    $\theta^{(x)}$. Black arrows correspond to the top-down generative model. Red arrows correspond to the bottom-up recognition model.}\label{graphicalModel}
\end{figure}

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', which draws samples from a recognition model (our $q$);
and  ``sleeping'', which trains a recognition model on samples from the generative model (our $\mathcal{D},\theta$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks.
The wake cycle can be thought of active problem solving,
guided by the current generative and recognition models.
A sleep cycle means  updating these models so that the DSL and  recognition model
more closely match the distribution of tasks.
An important difference from the Helmholtz machine
is that our setup has two distinct sleep phases: a phase where we update the generative model
$(\mathcal{D},\theta)$, which we call \textbf{Sleep-G},
and a second phase where we update the recognition model $q$,
which we call \textbf{Sleep-R}.
This wake/sleep style of inference is the reason why we call our approach \system.

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference in the generative model of Figure~\ref{graphicalModel}(a) by maximizing this lower bound on the joint probability,
alternatingly maximizing w.r.t. the frontiers (Wake) and the DSL (Sleep-G):
%So interleaved with the DSL induction steps are program induction steps. Section~\ref{explorationSection}
%explains how \system synthesizes new programs and given a DSL.
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by searching for programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

Searching for programs is hard because
of the large combinatorial search space. We ease this difficulty by learning a neural recognition model
during the Sleep-R phase:

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train a neural network, $q$, to predict a
distribution over programs conditioned on a task. The objective of $q$
is to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
The recognition model exploits the structure of the DSL: %$\mathcal{D}$:
rather than directly predicting a distribution over $p$ conditioned on $x$,
it predicts a distribution over components of the DSL.
This taps into the intuition that programming is primarily a top-down activity:
as human programmers, we first decide what kind of
programming constructs we might need to use,
and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Figure~\ref{graphicalModel}.

Our lower bound $\lowerBound$ is unconventional,
and one might wonder why we do not instead maximize an ELBO-style bound like in a VAE or in the EM algorithm.
We explain in the supplement that maximizing an ELBO-style bound
leads to a pathological behavior that  causes the model to easily become trapped in local optima.

@@

The recognition model $q$ is a neural network that predicts,
for each task $x\in X$, a weight vector $q(x) = \theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.
Together with the DSL, this defines a distribution over programs,
$\probability[p|\mathcal{D},\theta = q(x)]$.
We abbreviate this distribution as $q(p|x)$.
The crucial aspect of this framing is that the neural network
leverages the structure of the learned DSL,
so it is \emph{not} responsible for
generating programs wholesale.
We share this aspect with the DeepCoder system~\cite{balog2016deepcoder}.

@@

To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
a DSL in \system may be context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
In the Supplement we derive a tractable MAP estimator for $\theta$.
\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

@@

Algorithm~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
We note the following implementation details:
(1) We interleave wake cycles with each of the sleep cycles;
(2) on the first iteration, we do \emph{not} train the recognition model on samples from
the generative model because a DSL has not yet been learned --
we instead train the network to only maximize $\mathcal{L}_{\text{AE}}$.
(3) Because the frontiers can grow very large,
we only keep around the top $10^4$ programs $p$ in each frontier $\mathcal{F}_x$
with the highest likelihood $\probability[x,p|\mathcal{D},\theta]$.
(4) During both DSL induction and neural net training, we calculate $\text{score}(\mathcal{D}')$ and $\mathcal{L}_{\text{RM}}$
by only summing over the top $K$
programs in $\mathcal{F}_x$ -- we found that $K = 2$ sufficed.
(5) For added robustness, we enumerate programs from both the generative model and the recognition model.

@@

This experimental setup is an extension of one introduced in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
%% For this domain our recognition model is a simple multilayer perceptron 
%% whose inputs are the input/output examples.
\system learns to build these
circuits by creating new subroutines that correspond to
logic gates and multiplexers (Figure~\ref{multiplexer}).
We will show data showing that the neural recognition model
assists in building these circuits (Section~\ref{quantitative}),
suggesting that an agent can glean something about the structure of a circuit just by looking at its truth table.

@@

We take approaches from two different initial DSLs, \emph{base} and
\emph{rich}, and assess their performance. The base DSL starts with only
low-level primitives such as \code{mapi}, \code{reducei}, and
\code{if}, whereas the rich DSL includes some common structure that can be
built from these such as \code{all}, \code{filter}, \code{slice}, and
\code{index}.
Both are capable of solving all tasks supplied by our dataset.
See Supplement for more detail on these DSLs.

@@

  \\\midrule\\\multicolumn{6}{c}{\emph{Boolean Circuits}}\\\midrule
  \% solved &
  \textbf{84\%} &76\%& 63\%&62\%&62\%\\
  Solve time&  0.9s&  1.1s&1.2s&1.3s&1.3s

  \\\midrule\\\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&   \textbf{ 98\% }&94\%&0.7\%&0\%&2\% \\
  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
  MDL (nats)&   9.3 &11.0&2.3&--&4.6

  \\\midrule\\\multicolumn{6}{c}{\emph{String Editing}}\\\midrule
  \% solved&\textbf{99\%} &82\% &9\%&24\%&18\%\\
  Solve time&  0.9s&2.2s&2.7s&2.3s&2.4s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (base DSL)}}\\\midrule
  \% solved&\textbf{52\%} &\textbf{52\%} &40\%&44\%&42\%\\
  Solve time&  0.4s&0.7s&1.2s&0.9s&1.0s

  \\\midrule\\\multicolumn{6}{c}{\emph{List functions (rich DSL)}}\\\midrule
  \% solved&\textbf{86\%} &84\% &60\%&74\%&70\%\\
  Solve time&  0.8s&0.7s&1.0s&1.0s&1.1s
  \\\bottomrule
  \end{tabular}
\caption{\% solved: fraction solved with 5 sec timeout. Solve time: averaged over solved tasks. MDL: $-\expect\left[\probability[x|p] \right]$. For domains other than symbolic regression MDL is 0 nats. For symbolic regression MDL is a proxy for \# of continuous degrees of freedom.}\label{baselineComparisons}  \end{table}

@@

 \noindent \emph{Neural recognition models for program induction:} Much recent work in the ML community has
 focused on creating neural networks that regress from
 input/output examples to programs~\cite{devlin2017robustfill,devlin2017neural,menon2013machine,balog2016deepcoder}. This family of work is closest to the lesioned version of
 \system where $(\mathcal{D},\theta)$ is held fixed and $q$ is learned (RF/DC baseline).
 These neural networks are typically trained with strong supervision (i.e., with annotated ground-truth programs) on massive data sets (i.e., hundreds of millions~\cite{devlin2017robustfill}).
 In our work we consider a weakly-supervised setting where ground truth programs are not provided and
 the agent must learn from a few hundred examples.

@@

 \noindent \emph{Inventing new subroutines for program induction:}
 Several program induction algorithms, most prominently the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}, take as their goal to learn new, reusable subroutines that are shared in a multitask setting. We find this work inspiring and motivating,
 and extend it along two dimensions: (1) we propose a new algorithm for
 inducing reusable subroutines, based on Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing};
 and (2) we show how to combine these techniques with bottom-up neural recognition models.
 Other instances of this related idea are~\cite{DBLP:conf/icml/LiangJK10} and Schmidhuber's OOPS model~\cite{schmidhuber2004optimal}.

@@

 \section{Conclusion}
 We contributed a model, \system, that learns to program by
 bootstrapping a DSL with new domain-specific primitives that the algorithm itself discovers.
  Alongside this DSL, we train a bottom-up
 neural recognition model that learns how to efficiently deploy the
 DSL and hone in on the right programs.  We believe that this
 integration of top-down symbolic representation and bottom-up neural
 networks -- both of them learned -- is a useful general approach for
 getting program induction off the ground.
